---
title: 'HW1'
output:
  html_document:
    df_print: paged
---

# Yermolova Daria

```{r}
library(wooldridge)
library(stargazer)
library(dplyr)
library(corrplot)
library(car)
library(jtools)
library(tidyverse)
library(haven)
library(lmtest)
library(skedastic)
library(sandwich)
library(margins)
library(robustbase)
```


#9.2

#(1)
```{r}
rm(list = ls())
data('wage2')
reg <- lm(lwage ~ educ + exper + tenure + married + south + urban + black + KWW, data = wage2)
stargazer(reg,type = 'text')
```
The coef is 0.058, which is close to the estimate obtained in example.

#(2)
```{r}
reg2 <- lm(lwage ~ educ + exper + tenure + married + south + urban + black + KWW + IQ, data = wage2)
stargazer(reg2,type = 'text')
```
The coeff is 0.05, it decreased by 1%.

#(3)
```{r}
lht(reg2, c('IQ','KWW'), c(0,0))
```

IQ and KWW are statistically significant at 5% confidence level. 
p-value is small, so both IQ and KWW are jointly statistically significant.

# 9.3

#(1)
Grant is probably correlated with other factors which impact productivity.
There are only 2 variables in the model, so this factor is contained in u.

#(2)
```{r}
rm(list = ls())
data(jtrain)
reg1 <- lm(log(scrap) ~ grant, data = filter(jtrain, year == 1988))
stargazer(reg1, type = 'text')
```
The coef grant is not statistically different from zero.

#(3)
```{r}
reg4 <- lm(log(scrap) ~ grant + lscrap_1, subset(jtrain,year == 1988))
stargazer(reg4, type = 'text')
```
The effect of grant increased, became significant and negative. The impact of grant is -25%.
−.254/.147 = -1.73. 
−1.73 < −1.68
we reject H0 at the 5% level.

#(4)
```{r}
lht(reg4, 'lscrap_1 = 1')
```
p-value = 0.0003885, which means we can reject H0.

#(5)
```{r}

reg5 <- coeftest(reg4, vcov = sandwich)
stargazer(reg5,reg4, type = 'text', digits = 4)

-0.2540/0.1422

(0.8312-1)/0.0715

```
T-statistic of the first test is: 
1.786217 > 1.675285, (significant)
Second test lscrap_1 = 1:
t-statistic = -2.360839,(significant) against 2.007584 t-value for two-sided test on 5% significance level.

#9.8

# (1)
```{r}
rm(list = ls())
data('twoyear')
mean(twoyear$stotal)
sd(twoyear$stotal)
```

#(2)
```{r}
reg <- lm(stotal ~ jc, data = twoyear)
reg2 <- lm(stotal ~ univ, data = twoyear)
stargazer(reg,reg2,type = 'text')
```
Only univ is statistically significant.

#(3)
```{r}
reg3 <- lm(lwage ~ jc + univ + exper + stotal, data = twoyear)
stargazer(reg3, type = 'text')
lht(reg3, 'jc = univ')
```
p-value = 0.4205/2 = 0.21
We fail to reject H0.
Thus, the return of four-year colleges is greater.

#(4)
```{r}
twoyear$var <- twoyear$stotal * twoyear$stotal
reg4 <- lm(lwage ~ jc + univ + exper + stotal + var, data = twoyear)
lht(reg4, c('var'), c(0))
```
p-value = 0.686, we fail to reject H0, so we do not need the quadratic term.

#(5)
```{r}
twoyear$var1 <- twoyear$stotal * twoyear$univ
twoyear$var2 <- twoyear$stotal * twoyear$jc
reg5 <- lm(lwage ~ jc + univ + exper + stotal + var1 + var2, data = twoyear)
lht(reg5, c('var1', 'var2'), c(0,0))
```
The interaction terms are jointly insignificant.


#(6)
The model from part 3 is not needed since the variables are statistically insignificant. All the variables are significant in the model from part 2 

